---
title: "Data607-Major Assignment-Project3-Most Valued Data Science Skills"
author: "Leo Yi / David Blumensteil / Manolis Manoli / Vinayak Kamath"
date: "3/11/2020"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: TRUE
    toc_depth: 3    
    toc_float:
      collapsed: true
      smooth_scroll: true
      number_sections: true    
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(stringr)
library(knitr)
library(tidyr)
library(dplyr)
#install.packages("wordcloud")
#library("wordcloud")
#install.packages("RColorBrewer")
#library("RColorBrewer")
```

### Import

```{r}
#Loads the dataset
df_raw <- read.csv("https://raw.githubusercontent.com/davidblumenstiel/data/master/indeed_job_dataset.csv",stringsAsFactors = FALSE)

#Create working df object
df <- df_raw
```

### Soft Skills

```{r}
#keywords as 'keywords'
keywords<-unname(unlist(read.csv("https://raw.githubusercontent.com/chilleundso/DATA607/master/Project3/softskills.csv", stringsAsFactors = FALSE)))
#Iniates new columns for each keyword; also makes list index list
df[keywords] <- NA
keywordColIndex <- seq(length(df)-length(keywords)+1,length(df),1)
#Will determine if contains keywords; occurs (1), or doesnt occur(0)
#Coerces multiple occurances to 1
i = 0
for (desc in df$Description) {
  i = i + 1
  df[i,keywordColIndex] <-(as.integer(str_detect(desc,fixed(keywords, ignore_case = TRUE))))
}
#Gets rid of the descriptions
df$Description <- NULL
#Gather soft skills
df2 <- df[,c(1, 44:ncol(df))] %>%
  gather(skill, flag, -X) %>%
  filter(flag == 1) %>%
  select(id = X, skill) %>%
  mutate(skill_type = 'soft')
```


```{r}
#Gather soft skills
df2 <- df[,c(1, 44:ncol(df))] %>%
  gather(skill, flag, -X) %>%
  filter(flag == 1) %>%
  select(id = X, skill) %>%
  mutate(skill_type = 'soft')
```

### Hard Skills


#Gather soft skills
df2 <- df[,c(1, 44:ncol(df))] %>%
  gather(skill, flag, -X) %>%
  filter(flag == 1) %>%
  select(id = X, skill) %>%
  mutate(skill_type = 'soft')

```

### Hard Skills

```{r}
# List all hard skills and how often they were listed in a separate dataframe
for (i in 1:nrow(df_raw)) {
  slist <- str_extract_all(df$Skill[i], "'.+?'")
  temp_df <- data.frame(skill = slist[[1]])

  if (nrow(temp_df) > 0) {

    temp_df$id <- i - 1
    temp_df$skill <- as.character(temp_df$skill)
    temp_df <- select(temp_df, id, skill)

    if (i == 1) {
      sdf <- temp_df
    } else {
      sdf <- bind_rows(sdf, temp_df)
    }
  }
}

# Remove quotes and trim whitespace
sdf$skill <- str_replace_all(sdf$skill, "'", "")
sdf$skill <- str_replace_all(sdf$skill, '"', "")
sdf$skill <- str_trim(sdf$skill)

sdf$skill_type <- 'hard'

# Combine skills
skills <- bind_rows(df2, sdf)
```


### Job Title

# Data Tidying/ Transformation

```{r, warning = FALSE}
df <- df_raw

job.indeed.df  <- df %>%
  select(c(Id=X, (Job_Title)))

# Pattern Building:
pattern.analyst <- c('analyst','statistician','analysis','analytics')
pattern.engineer <- c('engineer', 'engg', 'technician','technologist','designer','architect')
pattern.scientist <- c('scientist','doctor','dr.')
pattern.junior <- c('junior','jr', 'entry','internship','jr.')
pattern.senior <- c('senior', 'sr','experienced','sr.')

# Intermedaite Data Frame for Titlles:
final.data.df <- data.frame(Id=integer(nrow(job.indeed.df)), Job_Title=character(nrow(job.indeed.df))
                            , analyst=integer(nrow(job.indeed.df)) ,engineer=integer(nrow(job.indeed.df)),scientist=integer(nrow(job.indeed.df))
                            , junior=integer(nrow(job.indeed.df)),senior=integer(nrow(job.indeed.df)))
final.data.df$Id <-   job.indeed.df$Id
final.data.df$Job_Title <- as.character( as.character( job.indeed.df$Job_Title) )

# Working on the counts:
for (i in 1: nrow(job.indeed.df)) {

  final.data.df$analyst[i] <- if(grepl(paste(pattern.analyst,collapse="|"), job.indeed.df$Job_Title[i], ignore.case = TRUE) )  1 else 0
  final.data.df$engineer[i] <- if(grepl(paste(pattern.engineer,collapse="|"), job.indeed.df$Job_Title[i], ignore.case = TRUE) )  1 else 0
  final.data.df$scientist[i] <- if(grepl(paste(pattern.scientist,collapse="|"), job.indeed.df$Job_Title[i], ignore.case = TRUE) )  1 else 0

  final.data.df$junior[i] <- if(grepl(paste(pattern.junior,collapse="|"), job.indeed.df$Job_Title[i], ignore.case = TRUE) )  1 else 0
  final.data.df$senior[i] <- if(grepl(paste(pattern.senior,collapse="|"), job.indeed.df$Job_Title[i], ignore.case = TRUE) )  1 else 0
}

#c.word <- c('analyst','engineer','scientist','junior','senior')
#c.count <- c(sum(final.data.df$analyst) ,sum(final.data.df$engineer), sum(final.data.df$scientist),sum(final.data.df$junior), sum(final.data.df$senior ) )
#set.seed(1234)
#wordcloud(words = c.word, freq = c.count, min.freq = 1,
#          max.words=10, random.order=FALSE, rot.per=0.35,
#          colors=brewer.pal(8, "Dark2"))
# Add job title flags to original dataframe

# Merging with main Analysis Data Frame:
df_all <- left_join(df_raw, final.data.df[,-2], by = c('X' = 'Id'))

```
